---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: yes
---


## Loading and preprocessing the data
```{r load, echo=TRUE}
mytable <- read.table("activity.csv", header = TRUE, sep = ",", stringsAsFactors = TRUE)
```

Here is a summary table
```{r mysummary, results='asis'}
library(knitr)
kable(summary(mytable))
```


## What is mean total number of steps taken per day?
I create a new dataframe variable called nonna without NAs. Here is a summary table

```{r nona, echo=TRUE}
nonna<-na.omit(mytable)
kable(summary(nonna))
```

Using dplyr package I calculate the total number of steps per day.
Then a histogram is plotted

```{r groupingdate, echo=TRUE}
library(dplyr)
steps_day <- nonna %>% 
  group_by(date) %>% 
  summarise(total=sum(steps), .groups= 'drop')
hist(steps_day$total,
     col = "peachpuff",
     border = "black",
     xlab = "Total number of steps per day",
     main = "Histogram: Total Number of Steps per Day")
```

Both measures are very close
```{r computeeda, echo=TRUE}
mymean <- mean(steps_day$total)
mymedian <- median(steps_day$total)
```

The mean of the total number of steps is `r mymean` and the median is `r mymedian`


## What is the average daily activity pattern?

Using dplyr package I calculate the mean of the number of steps taken.
Then a line chart is plotted

```{r groupinginterval, echo=TRUE}
steps_daily <- nonna %>% 
  group_by(interval) %>% 
  summarise(totalmean=mean(steps), .groups= 'drop')
plot(steps_daily$interval, steps_daily$totalmean, type = "l",
     col = 'red', lwd = 2,
     xlab = '5-Minute Interval', ylab = 'Average Number of Steps',
     main = 'Average Daily Activity Pattern')
```


Get the maximum number of steps from the 5-minute interval, on average across all the days in the dataset.


```{r computemax, echo=TRUE}
maxmean <- max(steps_daily$totalmean)
maxinterval <- steps_daily$interval[which.max(steps_daily$totalmean)]
```
The maximum average number of steps is `r maxmean` in the interval `r maxinterval`.


## Imputing missing values
Get the total number of missing values in the dataset
```{r sumnas, echo=TRUE}
count_nas <- sum(is.na(mytable))
```
The total number of missing values is `r count_nas`.

The imputation strategy chosen is the one based on the mean for that 5-minute interval

```{r meanperinterval, echo=TRUE}
meanvalues <-  nonna %>% 
  group_by(interval) %>% 
  summarise(totalmean=mean(steps), .groups= 'drop')
idx <- which(is.na(mytable))
cpytable <- mytable
for (i in 1:length(idx)){
    valueimputed<-meanvalues$totalmean[meanvalues$interval==cpytable$interval[idx[i]]]
    cpytable$steps[idx[i]] <-valueimputed
}
```


Here is a summary of the filled dataset
```{r summaryfilled, echo=TRUE}
kable(summary(cpytable))
```

Here is the histogram from the filled dataset
```{r histofilled, echo=TRUE}
steps_filled <- cpytable %>% 
  group_by(date) %>% 
  summarise(total=sum(steps), .groups= 'drop')
hist(steps_filled$total,
     col = "peachpuff",
     border = "black",
     xlab = "Total number of steps per day",
     main = "Histogram: Total Number of Steps per Day Filled Dataset")
```


```{r computefilled, echo=TRUE}
fillmymean <- mean(steps_filled$total)
fillmymedian <- median(steps_filled$total)
```

The mean and the median in the imputed dataset are the same `r fillmymean` and `r fillmymedian`, respectively. Compared against the original dataset (no NA's), the mean are the same (Original (No NA's)=`r mymean`, imputed = `r fillmymean`), but, the median is slightly different (Original (no NA's) = `r mymedian`, Imputed = `r fillmymedian`)


## Are there differences in activity patterns between weekdays and weekends?

```{r finde, echo=TRUE}
cpytable$day<-c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday","Friday", "Saturday")[as.POSIXlt(cpytable$date)$wday + 1]
cpytable$day[cpytable$day %in% c("Saturday", "Sunday")]<-"weekend"
cpytable$day[cpytable$day != "weekend"] <- "weekday"
wkd_table <- cpytable %>% filter(day == "weekday")
wknd_table <- cpytable %>% filter(day == "weekend")
cpytable <- aggregate(steps ~ interval + day, data=cpytable, mean)
library(lattice)
library(ggplot2)
ggplot(cpytable, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(day ~ .) +
  xlab("Interval") + 
  ylab("Number of Steps")+theme_bw()
```
